---
title: "PD_Schloss_NMDs_with_rarefaction"
output: html_notebook
---

# calculte distance with vegdist (vegan)
Here i'm following Schloss tutorial/ paper(https://doi.org/10.1101/2023.06.23.546313).
We can also see from my rarefaction curve that most of the sample are already in saturation in this seq point so I'm not losing a lot of data.
I also exclude the sampling from 2019 (the only sampling we didn't use separate tool kit for each tissue)
Before removing it, I preform the analysis with all of the sample as well
```{r}
library(tidyverse)
library(vegan)
library(ggplot2)
library(microbiomer)
library(phyloseq)
library(pairwiseAdonis)
library(ggpubr)
library(ARTool)
library(glue)
library(dplyr)
library(gridExtra)
```
# Upload phyloseq object 
```{r}
setwd("C:/Users/barak/OneDrive/Noga_Phd/Jellyfish_research/16S_analysis/Microbiome_jellyfish_project/processed_data")

# Load phyloseq 
biome_phyloseq<-readRDS("./bloom_phyloseq")
sample_data(biome_phyloseq)$Tissue_type <- ifelse(grepl("Bw", sample_data(biome_phyloseq)$Sample_name), "Bell wounded", sample_data(biome_phyloseq)$Tissue_type)
taxa_names(biome_phyloseq) <- paste0("ASV", seq(ntaxa(biome_phyloseq))) 

# Adding a column into the OTU table containing ASV_Genus column (useful in ASV graphic analysis)
asv_taxa_meta<-tax_table(biome_phyloseq) %>% 
         as.data.frame() %>% 
         rownames_to_column(var = "OTU") %>% 
         mutate(ASV_Genus = glue("{OTU}:{Genus}"))%>% 
         column_to_rownames(var = "OTU") %>% 
         tax_table() 
        
col<-c("Kingdom", "Phylum" ,"Class" , "Order"  ,"Family" ,"Genus",'ASV_Genus') 
row<-taxa_names(biome_phyloseq)

taxa_names(asv_taxa_meta)<-row
colnames(asv_taxa_meta@.Data)<-col

tax_table(biome_phyloseq)<-asv_taxa_meta  

#Only jellyfish samples 
phy_jelly <- biome_phyloseq %>% 
                   subset_samples(., Tissue_type %in% c("Bell","Tentacles","GVC","Gonads")&
                                     Year!="2019") %>% 
                   prune_taxa(taxa_sums(.) > 99, .) %>% 
                   prune_samples(sample_sums(.) > 14000, .)%>%  
                   pres_abund_filter(., pres = 2,abund =0, verbose = TRUE)

```

```{r}
# this is the ASVs table, i called it shared file to fit the tutorial I'm following 
metadata<- sample_data(phy_jelly) %>% 
           as_tibble(.) %>% 
           mutate(Sample_name= glue("{Sample_name}_{Month}"))

otu_table<-otu_table (phy_jelly) %>% 
            t() %>%  
            as.data.frame () %>% 
            mutate(Group=metadata$Sample_name) %>%
            pivot_longer(-Group) 

shared<- otu_table %>% 
            pivot_wider(Group) %>%
            as.data.frame()
rownames(shared)<-shared$Group

shared_matrix<-shared %>% 
        .[,-1] %>% 
        as.matrix(.)
```

```{r}
set.seed(220389)
dist_all<-avgdist(shared_matrix, dmethod="bray", sample=min(sample_sums(phy_jelly)),iterations=100)
nmds_all<-metaMDS(dist_all,k=3)

   stress_all<-nmds_all$stress %>%
      format(digits = 2)

    scores_all<- scores(nmds_all) %>%
      as.data.frame() %>%
      mutate(Sample_name=rownames(.))%>%
      right_join(.,metadata,by="Sample_name") %>%
      na.omit() %>%
      select(Sample_name,Tissue_type,Month,everything())

    centroid_all<- scores_all %>% 
    group_by(Tissue_type) %>% 
    summarize(NMDS1=mean(NMDS1),NMDS2=mean(NMDS2),NMDS3=mean(NMDS3)) %>% 
    as_data_frame()
```


```{r}
Tissue_color = c("Bell"="#0E4CAA","GVC"="green","Gonads"="purple", "Rhopalium"="orange","Tentacles"="gold4","Sea water"="deepskyblue3")

# Create a list of x and y pairs
xy_list <- list(one = c("NMDS1", "NMDS2"), two = c("NMDS1", "NMDS3"), three = c("NMDS2", "NMDS3"))

# Define the function to create the plot
create_plot <- function(xy) {
  x <- xy[1]
  y <- xy[2]
  ggplot(scores_all, aes_string(x = x, y = y ,color = "Tissue_type")) +
    geom_point(size=3,alpha=0.8)+
    stat_ellipse(aes(group = Tissue_type, color = Tissue_type),
                 linetype  = 2,
                 level=0.85,
                 show.legend = FALSE) +
    scale_color_manual(values = Tissue_color)+
    scale_fill_manual(values = Tissue_color)+
    annotate("text", x =.5, y = -.5, label = paste0("stress: ", stress_all), hjust = 0) +
    theme_bw()+
    theme(text = element_text(size = 12),
          legend.key.size = unit(4, 'mm'),
          strip.text = element_text(size = 12),
          legend.text = element_text(size = 12),
          strip.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.position = "right")+
    coord_cartesian(xlim = c(.65, -.65), ylim = c(.65, -.65))+
    labs(x = x, y = y, color = "Tissue type")
}

# Use lapply to apply the function to each pair in the list
plots <- lapply(xy_list, create_plot)
```

# Check statistic 
```{r}
adonis2(dist_all~scores(nmds_all))
pairwise.adonis(dist_all, scores_all$Sampling_date)
bd_bell <- betadisper(bell_dist, results_nmds[['B']]$scores$Month)
anova(bd_bell)
permutest(bd_bell)
TukeyHSD(bd_bell)
```

```{r}
ggsave("./nmds_all_month_3d.pdf", plot = all_month_3d, width = 16, height = 5)
ggsave("./NMDS/seawater_tissue_3d.pdf", plot = seawater_tissue_3d, width = 16, height = 5)
```
#calculte distance with vegdist (vegan) and bray curtis
```{r}
set.seed(220389)
tissue<-c("B","T","V","G")

results_nmds <- list()

for (i in tissue ){
    count_table <- otu_table %>%
      select(Group,value) %>%
      filter(grepl(i,Group)) %>%
      group_by(Group) %>%
      summarize(total = sum(value)) %>%
      arrange(total) %>%
      print(n=20)

    min_n_seq <- min(count_table$total)

    shared_sub <- shared[grepl(i, rownames(shared)),-1]

    dist<-avgdist(shared_sub, dmethod="bray", sample=min_n_seq,iterations =100)
    nmds<-metaMDS(dist)

    stress<-nmds$stress %>%
      format(digits = 2)

    scores<- scores(nmds) %>%
      as.data.frame() %>%
      mutate(Sample_name=rownames(.))%>%
      right_join(.,metadata,by="Sample_name") %>%
      na.omit() %>%
      select(Sample_name,Month,everything())

    centroid<- scores %>% 
    group_by(Month) %>% 
    summarize(NMDS1=mean(NMDS1),NMDS2=mean(NMDS2)) %>% 
    as_data_frame()
    
    
    # Store results for current tissue type
    results_nmds[[i]]<-list(min_n_seq=min_n_seq,
                          scores=scores,
                          stress=stress,
                          centroid=centroid,
                          dist=dist)
}
```
# Plot NMDS for each tissue_type
```{r}
# Define month colors
month_colors <- c(
  'Feb' = '#31cdd3ff',
  'Jun' = '#f54441ff',
  'Jul' = '#ddb63b'
)

set.seed(220389)

# Define tissue types
tissue <- c("B", "T", "V", "G")

# Helper function to create a plot for a given tissue type
create_nmds_plot <- function(tissue_type) {
  scores <- results_nmds[[tissue_type]]$scores %>% as_tibble()
  stress <- results_nmds[[tissue_type]]$stress
  centroid <- results_nmds[[tissue_type]]$centroid %>% as_tibble()

  ggplot(scores, aes(x = NMDS1, y = NMDS2, color = Month)) +
    geom_point(size = 3) +
    stat_ellipse(
      aes(group = Month),
      linetype = 2,
      level = 0.9,
      show.legend = FALSE
    ) +
    scale_color_manual(values = month_colors, limits = c('Feb', 'Jun', 'Jul')) +
    geom_point(
      data = centroid,
      color = "black",
      size = 4,
      shape = 21,
      aes(fill = Month), # Map fill to Month for centroids
      show.legend = FALSE
    ) +
    scale_fill_manual(values = month_colors) + # Add fill scale for centroids
    annotate("text", x = 1.05, y = -1, label = paste0("stress: ", stress), hjust = 0) +
    theme_bw(base_size = 12) +
    theme(
      legend.key.size = unit(4, 'mm'),
      strip.text = element_text(size = 12),
      strip.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.position = "right"
    ) +
    coord_cartesian(xlim = c(-1.05, 1.05), ylim = c(-1.05, 1.05)) +
    labs(x = "NMDS1", y = "NMDS2", color = "Month", title = paste0("Tissue Type: ", tissue_type))
}

# Generate plots for all tissue types
plots <- lapply(tissue, create_nmds_plot)

# View the plots in a grid
nmds_no_2019 <- grid.arrange(grobs = plots[c(1, 4, 3, 2)], nrow = 1)
```

```{r}
save.image("./NMDS.rds")
ggsave("./nmds_no_2019.pdf", plot = nmds_no_2019, width = 22, height = 5)
```

# Adonis (PERMANOVA) and betadisper
```{r}
bell_dist<-results_nmds[['B']]$dist
gonads_dist<-results_nmds[['G']]$dist
gvc_dist<-results_nmds[['V']]$dist
tentacles_dist<-results_nmds[['T']]$dist

adonis2(bell_dist~results_nmds[['B']]$scores$Month)
pairwise.adonis(bell_dist, results_nmds[['B']]$scores$Month)
bd_bell <- betadisper(bell_dist, results_nmds[['B']]$scores$Month)
anova(bd_bell)
permutest(bd_bell)
TukeyHSD(bd_bell)

adonis2(gonads_dist~results_nmds[['G']]$scores$Month)
pairwise.adonis(gonads_dist, results_nmds[['G']]$scores$Month)
bd_gonads <- betadisper(gonads_dist, results_nmds[['G']]$scores$Month)
anova(bd_gonads)
permutest(bd_gonads)
TukeyHSD(bd_gonads)

adonis2(gvc_dist~results_nmds[['V']]$scores$Month)
pairwise.adonis(gvc_dist, results_nmds[['V']]$scores$Month)
bd_gvc <- betadisper(gvc_dist, results_nmds[['V']]$scores$Month)
anova(bd_gvc)
permutest(bd_gvc)
TukeyHSD(bd_gvc)

adonis2(tentacles_dist~results_nmds[['T']]$scores$Month)
pairwise.adonis(tentacles_dist, results_nmds[['T']]$scores$Month)
bd_tentacles<- betadisper(tentacles_dist, results_nmds[['T']]$scores$Month)
anova(bd_tentacles)
permutest(bd_tentacles)
TukeyHSD(bd_tentacles)
```

#Jitter ordinations plot 
```{r}
dist_jitter_all<-dist_all%>% 
  as.matrix() %>% 
  as_tibble(rownames = "samples") %>% 
  pivot_longer(-samples) %>% 
  filter(samples < name)

  
jitter_tissue<-dist_jitter_all %>% 
      as.data.frame() %>%
      separate(samples, into=c("Sample_name_a", "Month_a"), "_", convert=TRUE) %>%
      separate(name, into=c("Sample_name_b", "Month_b"), "_", convert=TRUE) %>%
      mutate(Tissue_type_a= case_when(grepl("G",Sample_name_a)~ "Gonads",
                                      grepl("B",Sample_name_a)~ "Bell",
                                      grepl("V",Sample_name_a)~ "GVC",
                                      grepl("T",Sample_name_a)~ "Tentacles"),
             Tissue_type_b= case_when(grepl("G",Sample_name_b)~ "Gonads",
                                      grepl("B",Sample_name_b)~ "Bell",
                                      grepl("V",Sample_name_b)~ "GVC",
                                      grepl("T",Sample_name_b)~ "Tentacles"),
             comparison = case_when(
                                      Tissue_type_a == "Bell" &  Tissue_type_b == "Bell"  ~ "Bell",
                                      Tissue_type_a == "Gonads" &  Tissue_type_b == "Gonads"  ~ "Gonads",
                                      Tissue_type_a == "GVC" &  Tissue_type_b == "GVC"  ~ "GVC",
                                      Tissue_type_a == "Tentacles" &  Tissue_type_b == "Tentacles"  ~ "Tentacles",
    TRUE ~ NA_character_)) %>% 
  drop_na() 

jitter_month<-dist_jitter_all %>% 
      as.data.frame() %>%
      separate(samples, into=c("Sample_name_a", "Month_a"), "_", convert=TRUE) %>%
      separate(name, into=c("Sample_name_b", "Month_b"), "_", convert=TRUE) %>%
      mutate(comparison = case_when(
                                      Month_a == "Feb" &  Month_b == "Feb"  ~ "Feb",
                                      Month_a == "Jun" &  Month_b == "Jun"  ~ "Jun",
                                      Month_a == "Jul" &  Month_b == "Jul"  ~ "Jul",
    TRUE ~ NA_character_),
    comparison = factor(comparison, levels=c("Feb", "Jun", "Jul"))
    ) %>% 
  drop_na()
 

ggplot(jitter_month,aes(x=comparison, y=value,color = comparison)) +
               geom_jitter(width = 0.25, size = 2)+
               scale_color_manual(values = c('Feb'='#31cdd3ff',"Jun"="#f54441ff",'Jul'="#ddb63b"))+ 
               theme_classic()+
               stat_summary(fun.data=median_hilow, color="black", size=1,
                             fun.args = list(conf.int=0.50))+
               theme(text = element_text(size = 12),
                      legend.key.size = unit(4, 'mm'),
                      strip.text = element_text(size = 12),
                      legend.text = element_text(size = 12),
                      legend.position = "none") +
               labs(x=NULL, y="Bray-Curtis dissimilarity")
```


```{r}
jitter_tissue_long<-jitter_tissue %>%
                    select(comparison,value) %>% 
                    as.tibble() %>% 
                    mutate(comparison=as_factor(comparison))

jitter_month_long<-jitter_month %>%
                    select(comparison,value) %>% 
                    as.tibble() %>% 
                    mutate(comparison=as_factor(comparison))
                    
                    
model_ordi_all=art(value~comparison, jitter_month_long) #tissue type
anova(model_ordi_all)
pair_tissue = art.con(model_ordi_all, "comparison")
pair_tissue

```

```{r}
dist_list<-c("B","G","T","V")
jitter_dis<-list()

for (i in dist_list){
dist_jitter<-results_nmds[[i]]$dist%>% 
  as.matrix() %>% 
  as_tibble(rownames = "samples") %>% 
  pivot_longer(-samples) %>% 
  filter(samples < name)

  
jitter<-dist_jitter %>% 
      as.data.frame() %>%
      separate(samples, into=c("Sample_name_a", "Month_a"), "_", convert=TRUE) %>%
      separate(name, into=c("Sample_name_b", "Month_b"), "_", convert=TRUE) %>%
      mutate(comparison = case_when(
                                      Month_a == "Feb" &  Month_b == "Feb"  ~ "Feb",
                                      Month_a == "Jun" &  Month_b == "Jun"  ~ "Jun",
                                      Month_a == "Jul" &  Month_b == "Jul"  ~ "Jul",
    TRUE ~ NA_character_),
    comparison = factor(comparison, levels=c("Feb", "Jun", "Jul"))
    ) %>% 
  drop_na()

jitter_dis[[i]]<-jitter
}

```
````{r}
tissue<-c("B","G","V","T")
pair_month<-list()
for(i in tissue) {
  jitter_long<-jitter_dis[[i]] %>%
    select(comparison,value) %>% 
    as.tibble() %>% 
  mutate(comparison=as_factor(comparison))
  model_ordi_month=art(value~comparison, jitter_long) #tissue type
  anova= anova(model_ordi_month)
  pair_month_i = art.con(model_ordi_month, "comparison")

  pair_month[[i]]<-pair_month_i
}
```

```{r}
set.seed(220389)
# Define a list of tissue types to plot
tissue<-c("B","T","V","G")

plots_jitter <- list()

# Loop over tissue types and create a plot for each tissue type
for (i in tissue) {
   jitter_dis_i <- jitter_dis[[i]] %>% 
                 as_tibble()
  
p<- ggplot(jitter_dis_i,aes(x=comparison, y=value)) +
            geom_jitter(width = 0.25, size = 1, color = "grey")+
           # scale_color_manual(values = c('Feb'='#31cdd3ff',"Jun"="#f54441ff",'Jul'="#ddb63b"))+ 
            stat_summary(fun.data=median_hilow, color="maroon", size=1,
                         fun.args = list(conf.int=0.50)) +
            labs(x=NULL, y="Bray-Curtis distances") +
            scale_y_continuous(limits=c(0, 1), breaks=seq(0, 1, 0.1)) +
            theme_classic()+
            ggtitle(paste("Tissue Type:", i))

  # Store the plot for the current tissue type
  plots_jitter[[i]] <- p
}


# View the plots for each tissue type
Jitter_ordi_t_p<-grid.arrange(grobs = plots_jitter, nrow = 1)
```



