---
title: "Figure 3"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
library(microbiomer)
library(tidyverse)
library(glue)
library(purrr)
library(stringr)
library(ggpubr)
library(rstatix)
```

```{r}
setwd("C:/Users/barak/OneDrive/Noga_Phd/Jellyfish_research/16S_analysis/Microbiome_jellyfish_project/")
# Load phyloseq object
biome_phyloseq<-readRDS("./processed_data/bloom_phyloseq")
sample_data(biome_phyloseq)$Tissue_type <- ifelse(grepl("Bw", sample_data(biome_phyloseq)$Sample_name), "Bell wounded", sample_data(biome_phyloseq)$Tissue_type)

# All samples without controls 
phy_jelly<- biome_phyloseq %>%
                          subset_samples(.,Tissue_type %in% 
                                    c("Bell",
                                      "Tentacles",
                                      "GVC",
                                      "Gonads"))%>%  
                          pres_abund_filter(., pres = 2,abund =0, verbose = TRUE) 
taxa_names(phy_jelly) <- paste0("ASV", seq(ntaxa(phy_jelly)))

phy_jelly_no_19 <-phy_jelly %>%
                          subset_samples(.,
                                      Year!="2019") %>% 
                          pres_abund_filter(., pres = 2,abund =0, verbose = TRUE) 

sum(phy_jelly_no_19@sam_data$Month=="Feb")
sum(phy_jelly_no_19@sam_data$Month=="Jun")
sum(phy_jelly_no_19@sam_data$Month=="Jul")

sum(phy_jelly_no_19@sam_data$Tissue_type=="Bell")
sum(phy_jelly_no_19@sam_data$Tissue_type=="Tentacles")
sum(phy_jelly_no_19@sam_data$Tissue_type=="GVC")
sum(phy_jelly_no_19@sam_data$Tissue_type=="Gonads")
```

```{r}
asv_counts<- psmelt(phy_jelly) %>%
  select(Sample_name,OTU)

asv_taxa_meta<-psmelt(phy_jelly) %>%
        mutate(Genus_ASV = glue("{OTU}:{Genus}"))

asv_taxa_meta_no_19<- asv_taxa_meta %>% 
                          subset(Year!= "2019")
```

#Transform the sample counts to relative abundance, calculte mean relative abundance, pool all taxa with relative abundance of <1%, plot at the ASV level
## Without 2019
2019 was the only year we didn't collected samples with separate tool kit, as it more prone to error I wanted to do analysis without 2019:July)

```{r}
colors_genus_otu<-c( "ASV1:Endozoicomonas" = "#930CED",                                  
                    "ASV2:Rickettsiales_uncl" = "#6B91E6",
                    "ASV3:type_III_uncl" = "#4B0920",
                    "ASV4:Simkaniaceae_uncl" = "#9A4516",
                    "ASV5:Tenacibaculum" = "#8DCA0B",
                    'ASV23:Flavobacteriaceae_uncl'="#628D07",
                    "ASV6:Bacteroides" = "#364E04",
                    "ASV13:Bacteroides" = "#A5B582",
                    "ASV7:Variovorax" = "#7f56b0",
                    "ASV9:Gammaproteobacteria_uncl" = "#4A007E",
                    "ASV15:Serratia" = "#7B208B",
                    "ASV10:Vibrio" = "#C99EF2",
                    "ASV25:Vibrio" = "#B694D4",
                    "ASV69:Vibrio" = "#C99EF2",
                    "ASV71:Vibrio"= "#C99EF2",
                    "ASV11:Mycoplasma" = "#FF4D6D",
                    "ASV21:Mycoplasma" = "#FCA4B4",
                    "ASV12:Helicobacteraceae_uncl" = "#D62828",
                    "ASV20:Helicobacteraceae_uncl" = "#CF6D6D",
                    "ASV14:Alkalibacterium" = "#9E174D",
                    "ASV16:Candidatus_Hepatoplasma" = "#FFAB5C",
                    "ASV18:Endozoicomonas" = "#B682D9",
                    "ASV19:Shewanella" = "gold4",
                    "ASV23:Fulvibacter" = "#579D1C",
                    "ASV24:Aquabacterium" = "#E7F48D",
                    "ASV26:Enhydrobacter" = "#DCFFFF",
                    "ASV27:Terasakiellaceae_uncl" = "#3B70A1",
                    "ASV29:Alkalibacterium" = "#E599C4",
                    "ASV37:Methylobacterium-Methylorubrum" = "#18874B",
                    "ASV39:Bradymonadaceae_uncl" = "#85E296",
                    "ASV72:Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium" = "yellow4",
                    "ASV87:Tenacibaculum" = "#A1CE42",
                    "ASV88:Psychrobium" = "#5A349E",
                    "ASV116:Aliivibrio" = "#FFD31F",
                    "ASV210:Alteromonadales_uncl" = "#D8B4FF",
                    "Other taxa" = "#A9A9A9")

```

## All samples divided to months 
```{r include=FALSE}
# For all samples
# Reshape the data from wide to long format
asv_rel_abund_no_19 <- asv_taxa_meta_no_19 %>% 
  group_by(Sample_name) %>%
  mutate(rel_abund = Abundance / sum(Abundance) * 100) %>%
  ungroup() %>%
  select(-Abundance) %>%
  pivot_longer(c("Phylum", "Class", "Order", "Family", "Genus", "OTU", "Genus_ASV"),
               names_to = "level",
               values_to = "taxon") %>% 
  mutate(Tissue_type = factor(Tissue_type),
         Month = factor(Month)) %>% 
  select(Sample_name, rel_abund, month_year, Tissue_type, Month, level, taxon)

mean_rel_abund_genus_no_19 <- asv_rel_abund_no_19 %>%
  filter(level == "Genus_ASV") %>%
  group_by(Month,Tissue_type, taxon) %>%
  summarize(mean_rel_abund = mean(rel_abund), .groups = "drop")

pool_genus_no_19 <- mean_rel_abund_genus_no_19 %>%
  group_by(Month, Tissue_type,taxon) %>%
  summarize(pool = max(mean_rel_abund) < 1,
            mean = mean(mean_rel_abund),
            max = max(mean_rel_abund),
            .groups = "drop")

Without_2019 <- inner_join(mean_rel_abund_genus_no_19, pool_genus_no_19, by = c("Tissue_type", "Month", "taxon"))

data_no_2019 <- Without_2019 %>%
  mutate(taxon = if_else(pool, "Other taxa", taxon)) %>%
  mutate(ASV_number = if_else(taxon == "Other taxa", Inf, as.numeric(str_extract(taxon, "\\d+"))),
         taxon = factor(taxon),
         Month = fct_relevel(Month, "Feb", "Jun", "Jul")) %>%
  group_by(Month,Tissue_type) %>%
  mutate(mean_rel_abund = mean_rel_abund / sum(mean_rel_abund) * 100) %>% # Renormalization step
  ungroup() %>%
  mutate(taxon = factor(taxon),
         taxon = fct_reorder(taxon, mean_rel_abund, .desc = FALSE),
         taxon = fct_shift(taxon, n = 1),
         taxon = fct_relevel(taxon, "ASV1:Endozoicomonas",
                                   "ASV18:Endozoicomonas",
                                   "ASV2:Rickettsiales_uncl",
                                   "ASV3:type_III_uncl",
                                   "ASV4:Simkaniaceae_uncl",
                                   "ASV5:Tenacibaculum",
                                   "ASV6:Bacteroides",
                                   "ASV13:Bacteroides",
                                   "ASV7:Variovorax",
                                   "ASV9:Gammaproteobacteria_uncl",
                                   "ASV11:Mycoplasma",
                                   "ASV21:Mycoplasma",
                                   "ASV12:Helicobacteraceae_uncl",
                                   "ASV14:Alkalibacterium",
                                   "ASV15:Serratia",
                                   "ASV16:Candidatus_Hepatoplasma",
                                   "ASV19:Shewanella",
                                   "ASV24:Aquabacterium",
                                   "ASV25:Vibrio",
                                   "ASV26:Enhydrobacter",
                                   "ASV29:Alkalibacterium",
                                   "ASV39:Bradymonadaceae_uncl",
                                   "Other taxa"))

facet_labels <- c('Bell' = "Bell (n=29)",
                  'Gonads' = "Gonads (n=31)",
                  'GVC' = "GVC (n=31)",
                  'Tentacles' = "Tentacles (n=31)")
# Plotting
ggplot(data_no_2019, aes(x = Month, y = mean_rel_abund, fill = taxon)) +
  geom_col() +
  facet_wrap(~Tissue_type, nrow = 1, labeller = labeller(Tissue_type = facet_labels))+
  scale_fill_manual(name = NULL, values = colors_genus_otu) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL,
       y = "Mean Relative Sequence Abundance (%)") +
  theme_classic() +
  theme(legend.position = "bottom",
        strip.background = element_blank(),       
        strip.text = element_text(face = "bold"))
```

```{r}
pdf("./bar.pdf",width = 5.2,height = 4.2)
ggplot(data_no_2019, aes(x = Month, y = mean_rel_abund, fill = taxon)) +
  geom_col() +
  facet_wrap(~Tissue_type, nrow = 1, labeller = labeller(Tissue_type = facet_labels))+
  scale_fill_manual(name = NULL, values = colors_genus_otu) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL,
       y = "Mean Relative Sequence Abundance (%)") +
  theme_classic() +
  theme(legend.position = "bottom",
        strip.background = element_blank(),       
        strip.text = element_text(face = "bold"))
dev.off()
```

## All tissues relative abundance 
```{r}
prepare_data_for_plot <- function(tissue) {
  
    tissue_bar <- asv_rel_abund_no_19 %>%
        filter(level %in% c("Genus_ASV"), Tissue_type == tissue) %>%
        group_by(Sample_name, month_year,Month, taxon, Tissue_type)
      
      pool_tissue_bar <- tissue_bar %>%
        group_by(Sample_name, taxon) %>%
        summarize(pool = max(rel_abund) < 2,
                  max = max(rel_abund),
                  .groups = "drop")
      
      tissue_data <- inner_join(tissue_bar, pool_tissue_bar, by = c("taxon", "Sample_name")) %>%
        mutate(taxon = if_else(pool, "Other taxa", taxon))
      
     # Create a data frame that maps each 'taxon' to its 'ASV_number'
    taxon_order <- tissue_data %>%
        filter(level %in% c("Genus_ASV"))%>%
        mutate(ASV_number = if_else(str_detect(taxon, "ASV"), as.numeric(str_extract(taxon, "\\d+")), Inf)) %>%
        select(taxon, ASV_number) %>%
        distinct() %>%
        arrange(ASV_number) %>% 
        distinct(taxon)
    
    # Convert the 'taxon' column to a vector
    taxon_vector <- taxon_order$taxon
    
    # Remove duplicates from the vector
    taxon_vector <- unique(taxon_vector)
     
     tissue_data <- inner_join(tissue_bar, pool_tissue_bar, by = c("taxon", "Sample_name")) %>%
        mutate(taxon = if_else(pool, "Other taxa", taxon)) %>%
        mutate(ASV_number = if_else(taxon == "Other taxa", Inf, as.numeric(str_extract(taxon, "\\d+"))),
               taxon = factor(taxon),
               taxon = fct_relevel(taxon, taxon_vector),
               Month = fct_relevel(Month, "Feb", "Jun", "Jul"))
  
  return(tissue_data)
}

# Apply the function to each tissue type
tissues <- c("Bell", "Gonads", "GVC", "Tentacles")

data_for_plots <- lapply(tissues, prepare_data_for_plot)
```

## plot all tissues relative abundance
```{r}
plot_list <- list()
for (i in seq_along(data_for_plots)) {
  # Create a plot for the current element
  plot <- ggplot(data_for_plots[[i]], aes(x=Sample_name, y=rel_abund, fill=taxon)) +
    facet_wrap(~Month, scales = "free_x", nrow=1) +
    geom_col() +
    scale_fill_manual(name=NULL, values=colors_genus_otu) +
    scale_y_continuous(expand=c(0, 0)) +
    labs(x=NULL, y="Relative Abundance (%)") +
    ggtitle(tissues[i]) +
    theme_minimal()+
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
          axis.text.y = element_text(size = 8),
          axis.title = element_text(),
          legend.position = "right",
          legend.title = element_text(size = 8)) 
  
  # Add the plot to your list
  plot_list[[i]] <- plot
}
plot_all<-ggarrange(plotlist = plot_list, ncol = 2, nrow = 2, common.legend = F,font.label=list(size=8), align = "hv", legend = "bottom")
```

```{r}
ggsave(filename="./rel_abund_all_samples.pdf", plot = plot_all, width = 14, height = 10,dpi = 300)
```


