---
title: "Top 25 ASVs from R.nomadica vs Seawater samples"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(phyloseq)
library(ggpubr)
library(glue)
library(microbiomer)
library(dplyr)
library(pheatmap)
library(stringr)
```

#### Upload R.nomadica phyloseq
```{r}
setwd("C:/Users/barak/OneDrive/Noga_Phd/Jellyfish_research/16S_analysis/Microbiome_jellyfish_project/processed_data/")

# Load phyloseq object
biome_phyloseq<-readRDS("./bloom_phyloseq")
sample_data(biome_phyloseq)$Tissue_type <- ifelse(grepl("Bw", sample_data(biome_phyloseq)$Sample_name), "Bell wounded", sample_data(biome_phyloseq)$Tissue_type)

# All samples without controls 
taxa_names(biome_phyloseq) <- paste0("ASV", seq(ntaxa(biome_phyloseq))) 

# Adding a column into the OTU table containing ASV_Genus column (useful in ASV graphic analysis)
add_otu_genus <- function(ps){
                  asv_taxa_meta<-tax_table(ps) %>% 
                           as.data.frame() %>% 
                           rownames_to_column(var = "OTU") %>% 
                           mutate(ASV_Genus = glue("{OTU}:{Genus}"))%>% 
                           column_to_rownames(var = "OTU") %>% 
                           tax_table() 
                          
                  col<-c("Kingdom", "Phylum" ,"Class" , "Order"  ,"Family" ,"Genus",'ASV_Genus') 
                  row<-taxa_names(ps)
                  
                  taxa_names(asv_taxa_meta)<-row
                  colnames(asv_taxa_meta@.Data)<-col
                  
                  tax_table(ps)<-asv_taxa_meta 
                  return(ps)
                }

biome_phyloseq <- add_otu_genus(biome_phyloseq)

#Only jellyfish samples 
rhopilema_ps<- biome_phyloseq %>% 
                   subset_samples(., Tissue_type %in% c("Bell","Tentacles","GVC","Gonads"),
                                     Year!="2019") %>% 
                   prune_taxa(taxa_sums(.) > 99, .) %>% 
                   prune_samples(sample_sums(.) > 14000, .)%>%  
                   pres_abund_filter(., pres = 2,abund =0, verbose = TRUE)

# Compute taxa sums
rhopilema_sum <- taxa_sums(rhopilema_ps) %>% 
                 sort(., decreasing = TRUE)

# Extract names of the 30 most abundant taxa
top_25_taxa <- names(rhopilema_sum)[1:25]

#extract the top 30 ASVs from rhopilema to the BLAST+ analysis 
top_25_rhopilema_ps <- prune_taxa(top_25_taxa,rhopilema_ps)
``` 
### Upload sea water 
# upload sea water data (themo and bloom)
```{r}
seawater_bloom_ps <- biome_phyloseq %>% 
                   subset_samples(., Tissue_type %in% c("Sea water")&
                                     Year!="2019") %>%
                   prune_samples(sample_sums(.) > 10000, .) %>% 
                   prune_taxa(taxa_sums(.) > 0, .)

seawater_bloom_ps <- add_otu_genus(seawater_bloom_ps)

setwd("C:/Users/barak/OneDrive/Noga_Phd/Jellyfish_research/16S_analysis/Jen_Project")

phyloseq_jen<-readRDS("./jen_phyloseq")


#Subset to only Themo samples (exclude ID 95 because it messed up the NMDS)
themo_ps<- subset_samples(phyloseq_jen,Cruise_Timepoint %in% c(1:12))%>%
                   prune_samples(sample_sums(.) > 5000, .) %>% 
                   prune_taxa(taxa_sums(.) > 50, .)


taxa_names(themo_ps) <- paste0("ASV", seq(ntaxa(themo_ps))) 


themo_ps <- add_otu_genus(themo_ps)
                   
```

Create ASV_Genus column with more info
```{r}
# Assuming you have a list of phyloseq objects named physeq_list
physeq_list <- list(rhopilema_ps,top_25_rhopilema_ps,seawater_bloom_ps,themo_ps)

# Loop through each phyloseq object
for (i in seq_along(physeq_list)) {
  # Extract tax_table and perform modifications
  asv_taxa_meta <- tax_table(physeq_list[[i]]) %>%
    as.data.frame()

  # Check if 'Species' column exists in the tax_table
  if ('Species' %in% colnames(asv_taxa_meta)) {
    # Exclude 'Species' column from the modifications
    asv_taxa_meta <- asv_taxa_meta %>%
      subset(select = -Species)
  }

  # Continue with the modifications using the pipe operator
  asv_taxa_meta <- asv_taxa_meta %>%
    rownames_to_column(var = "OTU") %>%
    mutate(ASV_dataset = glue("{OTU}:{Genus}:{i}")) %>%
    column_to_rownames(var = "OTU") %>%
    tax_table()

  # Update tax_table in the original phyloseq object
  col <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", 'ASV_Genus', 'ASV_dataset')
  row <- taxa_names(physeq_list[[i]])

  taxa_names(asv_taxa_meta) <- row
  colnames(asv_taxa_meta@.Data) <- col

  tax_table(tax_table(physeq_list[[i]])) <- asv_taxa_meta
}
```
# Write FASTA from phyloseq
```{r}
# Define the output file names
output_files <- c("asvs_rhopilema.fasta","asvs_cassiopea.fasta", "asvs_tri_mas.fasta","asvs_scyp.fasta", "asvs_aurelia.fasta","top_25_asvs_rhopilema.fasta","asvs_seawater_bloom.fasta","asvs_themo.fasta",'asvs_pulmo.fasta')

# Define the indices of the physeq_list elements you want to process
indices <- c(1, 2, 3, 4, 5,6,7,8,9)

# Loop over each index
for (i in seq_along(indices)) {
  # Get the current index
  index <- indices[i]
  
  # Get the current output file
  output_file <- output_files[i]
  
  # Create the full path to the output file
  output_path <- paste0("//wsl.localhost/Ubuntu-22.04/home/noga/Jellfish_16S/ncbi_jellyfish_comparison/", output_file)
  
  # Apply the ps_to_fasta function
  ps_to_fasta(physeq_list[[index]], out.file = output_path, seqnames = "ASV_dataset", width = 1000)
}

# from this stage I'm moving to the linux to preform a BLast+ comprison for each object vs the r.nomadica top 25 taxa.
```

# Upload BLAST+ results R,nomadica dominant only
```{r}
# Set the file path
setwd ("//wsl.localhost/Ubuntu-22.04/home/noga/Jellfish_16S/ncbi_jellyfish_comparison/")

# Read the file into a data frame
dominant_themo_blast <- read.table("./themo_vs_dominant_nomadica.txt", header = F, sep = "\t")
dominant_seawater_bloom_blast <- read.table("./seawater_bloom_vs_dominant_nomadica.txt", header = F, sep = "\t")


# Define Blast headers
blast_headers <- c(
  "Percentage_Identity",
  "Alignment_Length",
  "Mismatches",
  "Gap_Opens",
  "Q_Start",
  "Q_End",
  "S_Start",
  "S_End",
  "E_Value",
  "Bit_Score"
)
# Assign names to the columns
colnames(dominant_themo_blast) <- c("Themo", "rhopilema",blast_headers)
colnames(dominant_seawater_bloom_blast) <- c("Seawater_bloom", "rhopilema",blast_headers) 

# Define a function to filter out BLAST resluts
filter_data <- function(df, asv_col, min_length) {
  df %>%
    # Extract the number after "ASV" in the ASV column and convert it to numeric
    mutate(asv_num = as.numeric(str_extract(!!sym(asv_col), "\\d+"))) %>%
    subset(Alignment_Length > min_length) %>%
    group_by(rhopilema) %>%
    # Keep only the rows where Percentage_Identity is equal to the maximum Percentage_Identity in the group
    filter(Percentage_Identity == max(Percentage_Identity)) %>%
    # If there are ties in Percentage_Identity, keep only the row where asv_num is the minimum
    filter(asv_num == min(asv_num)) %>%
    ungroup() %>% 
    select(-asv_num) %>% 
    mutate(rhopilema_num = as.numeric(str_extract(rhopilema, "\\d+"))) %>% 
    arrange(rhopilema_num) %>% 
    select(rhopilema,!!sym(asv_col),Percentage_Identity) %>% 
    # Rename the Percentage_Identity column
    rename(!!paste("Percentage_Identity", asv_col, sep = "_") := Percentage_Identity)
}

# Apply the function to each data frame
themo_blast_filt <- filter_data(dominant_themo_blast, "Themo", 200)
seawater_bloom_blast_filt <- filter_data(dominant_seawater_bloom_blast, "Seawater_bloom", 200)
```

# Create merge df with all jellyfish ASVs, divide data to all jellyfish type
```{r}
# Merge the data frames by rhopilema
merged_data <- merge(seawater_bloom_blast_filt, themo_blast_filt, by = "rhopilema", all = TRUE) 

merged_data <- merged_data %>%
  mutate(Rhopilema_nomadica = rhopilema,
         Percentage_Identity_Rhopilema_nomadica= 100)%>% 
         select (-rhopilema)

# Convert "rhopilema" column to numeric
merged_data$rhopilema_numeric <- as.numeric(gsub("ASV(\\d+).*", "\\1", merged_data$Rhopilema_nomadica), na.rm = TRUE)

rhopilema_seqs <- sub(":.*", "", merged_data$Rhopilema_nomadica[!is.na(merged_data$Rhopilema_nomadica)])
themo_seqs <- sub(":.*", "", merged_data$Themo[!is.na(merged_data$Themo)])
seawater_bloom_seqs <- sub(":.*", "", merged_data$Seawater_bloom[!is.na(merged_data$Seawater_bloom)])
```

# Transform all phyloseq to relative abundance
```{r}
rhopilema_ps_rel<- transform_sample_counts(rhopilema_ps,function(x) x / sum(x)*100)
themo_ps_rel<- transform_sample_counts(themo_ps,function(x) x / sum(x)*100)
seawater_bloom_ps_rel<- transform_sample_counts(seawater_bloom_ps,function(x) x / sum(x)*100)
```

# Prune all ASVs to contain only the shared ASVs
```{r}
rhopilema_ps_shared<-prune_taxa(rhopilema_seqs,rhopilema_ps_rel)
themo_ps_shared<-prune_taxa(themo_seqs,themo_ps_rel)
seawater_bloom_ps_shared<-prune_taxa(seawater_bloom_seqs,seawater_bloom_ps_rel)


Rhopilema_nomadica<-psmelt(rhopilema_ps_shared) %>%
  mutate(jellyfish_type = str_replace(jellyfish_type, " ", "_"))

Themo <- psmelt(themo_ps_shared) %>% 
         mutate(jellyfish_type = "Themo")

Seawater_bloom <- psmelt(seawater_bloom_ps_shared)%>% 
                  mutate(jellyfish_type = "Seawater_bloom")

```
# Create a distinct taxa df for each jellyfish. calculate mean and max of each taxa relative abundance
```{r}
jellyfish_list <- c("Rhopilema_nomadica","Themo","Seawater_bloom")

for (obj_name in jellyfish_list) {
  obj <- get(obj_name)  # Get the actual object using its name
  
  distinct_max <- obj %>% 
    pivot_longer(c('OTU'),
                 names_to = "level",
                 values_to = "taxon") %>% 
    group_by(taxon) %>% 
    mutate(max = max(Abundance),
           mean = mean(Abundance)) %>% 
    distinct(jellyfish_type, Abundance, max, mean, taxon)
  
  distinct <- distinct_max %>%
    distinct(taxon, .keep_all = TRUE)
  
  # Update the original object with the modified data
  assign(paste0("distinct_",obj_name), distinct)
}
```

# Create new object to pull the relative abundance into
```{r}
merged_data_abundance <- merged_data %>% 
                         as_tibble() %>% 
                         mutate(ASVs = Rhopilema_nomadica,
                                rhopilema_numeric=as.numeric(rhopilema_numeric)) %>% 
                         mutate_at(vars(-c(ASVs,rhopilema_numeric)), function(x) sub(":.*$", "", x)) %>% 
                         mutate(ASVs = sub(":6", "", ASVs)) %>% 
                         select(rhopilema_numeric,ASVs,everything())
```

# Loop over all jellyfish type, pull relative abundance for eace jellyfish type. 
```{r}
calculate_abundance <- function(data, jellyfish_list, value) {
  for (i in jellyfish_list) {
    jellyfish_df <- get(paste0("distinct_", i))
    
    # Convert the i column in data to character
    data[[i]] <- as.character(data[[i]])
    
    joined_data <- left_join(data, jellyfish_df %>% select(taxon, !!value), by = setNames("taxon", i))
    data <- joined_data %>%
      mutate(!!sym(i) := ifelse(!is.na(!!sym(value)), !!sym(value), !!sym(i))) %>%
      select(-!!value)
  }
  return(data)
}

# Calculate max and mean abundances
merged_data_max_abundance <- calculate_abundance(merged_data_abundance, jellyfish_list, "max")
merged_data_mean_abundance <- calculate_abundance(merged_data_abundance, jellyfish_list, "mean")
```

# Orginized merged_data_abundance to fit ggplot bubble plot
```{r}
# Function for further processing
process_data <- function(data) {
  data[, -c(1:2)] <- lapply(data[, -c(1:2)], as.numeric)
  data <- data %>%
    mutate_at(vars(-ASVs), ~ ifelse(str_detect(., "ASV"), NA, .)) %>% 
    arrange(rhopilema_numeric) 
  
  data$ASVs <- sub(":1$", "", data$ASVs)
  # Replace all NA values with "0"
  data[is.na(data)] <- 0
  # Separate the columns that start with "Percentage_Identity" and those that don't
  cols <- names(data)
  Percentage_Identity <- cols[startsWith(cols, "Percentage_Identity")]
  other_cols <- cols[!startsWith(cols, "Percentage_Identity")]
  # Combine them back in the desired order
  new_order_cols <- c(other_cols, Percentage_Identity)
  # Reorder the dataframe
  data <- data[, new_order_cols]
  return(data)
}

# Apply the function to both data frames
merged_data_max_abundance <- process_data(merged_data_max_abundance)
merged_data_mean_abundance <- process_data(merged_data_mean_abundance)

# Species_col
Species_col <- names(merged_data_mean_abundance)[3:5]
```

```{r}
process_final_data <- function(data) {
  # Reshape the data to a longer format
  long_data <- merged_data_mean_abundance %>%
    pivot_longer(
      cols = Species_col,
      names_to = "Species",
      values_to = "Relative_abundance"
    )

  reshaped_data <- long_data %>%
    pivot_longer(
      cols = starts_with("Percentage_Identity_"),
      names_to = "Percentage_Identity",
      values_to = "Percentage"
    ) %>%
    mutate(Percentage_Identity = str_remove(Percentage_Identity, "Percentage_Identity_"))

  # Define the breaks for the categories
  breaks <- c(-Inf,97, 98, 99, 100, Inf)

  # Define the labels for the categories
  labels <- c("<97", "97", "98", "99", "100")

  final_data <- reshaped_data %>%
    filter(Relative_abundance != 0) %>% 
    mutate(Percentage_correct = ifelse(Species == Percentage_Identity, Percentage, NA),
           Percentage_correct = as.numeric(Percentage_correct),
           log_Relative_abundance = log1p(Relative_abundance))

 final_data <- final_data %>% 
  mutate(
    Percentage_Category = cut(final_data$Percentage_correct, breaks = breaks, labels = labels, right = FALSE),
    Species = factor(Species, levels = c("Themo",
                                         "Seawater_bloom",
                                         "Rhopilema_nomadica")),
    ASVs = factor(ASVs, levels = rev(unique(ASVs[order(rhopilema_numeric)])))) %>% 
  group_by(ASVs, Species) %>%
  fill(c(Percentage_correct), .direction = "downup") %>%
  ungroup() %>% 
  filter(Percentage_Category != '<97')%>%
  select(rhopilema_numeric,ASVs,Species,Percentage_Identity,Relative_abundance, log_Relative_abundance,everything(),-Percentage) %>% 
  distinct(ASVs, Species, .keep_all = TRUE)

return(final_data)
}
# Apply the function to both data frames
final_data_mean_abundance <- process_final_data(merged_data_mean_abundance)
final_data_max_abundance <- process_final_data(merged_data_max_abundance)

# Define the colors for the categories
colors_jelly <- c("<97" = "darkgrey", "97" = "#d1bdde", "98" = "#af74d6",'99' = '#930CED',"100" = "#440154ff")
break_size <- c(0.01,0.1,1, 10,20,30,47)
```

```{r}
# Filter rows where Species is 'Rhopilema_nomadica'
rhopilema_data <- final_data_mean_abundance %>%
  filter(Species == "Rhopilema_nomadica")

# Order the ASVs based on the relative abundance of Cassiopea_xamachana
ordered_ASVs <- rhopilema_data %>%
  arrange(Relative_abundance) %>%
  pull(ASVs)


# Create the bubble plot
ggplot(final_data_mean_abundance, aes(y = factor(ASVs, levels = ordered_ASVs), x = Species, size = Relative_abundance, color = Percentage_Category)) +
  # Add blue background for specific columns using annotate
  annotate("rect", xmin = 0.5, xmax = 2.5, ymin = -Inf, ymax = Inf, fill = "lightblue", alpha = 0.5) +
  # Add the original points
  geom_point(alpha = 1) +
  scale_size(range = c(1, 12), breaks = break_size) +
  scale_color_manual(values = colors_jelly) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 410, hjust = 1, size = 9),
        axis.text.y = element_text(size = 9),
        legend.position = "right",
        legend.title = element_text(size = 9)) +
  guides(color = guide_legend(override.aes = list(size = 10))) +
  labs(y = "Top ASVs R.nomadica")
``` 


